### 2022-09-28 procedure/code/203_results.Rmd
Create 203_results.Rmd
add markdown for visualizing results

---
title: "Reproduction Survey Analysis"
author: "Joseph Holler, Sarah Bardin, Peter Kedron"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(here)
library(tidyverse)
```

## R Markdown

Abstract...

```{r load}
survey_resp <- readRDS(here("data", "derived", "public", "survey_resp.RDS"))
summary(survey_resp)
```

```{r sub-fields-by-methods}
# crosstab sub-fields and methods
ftable(data = survey_resp, Q4 ~ Q3)
```

```{r familiarity-by-subfield-and-method}
# crosstab familiarity with sub-fields and methods
ftable(data = survey_resp, Q5 ~ Q3 + Q4)
```

```{r familiarity-by-subfield}

# calculate familiarity with reproducible practices
survey_resp <- survey_resp %>% mutate(familiar = as.numeric(Q7a) + 
                                        as.numeric(Q7b) + as.numeric(Q7c) +
                                        as.numeric(Q7d) + as.numeric(Q7e))

# summarize and ANOVA familiarity with practices by subfield
group_by(survey_resp, Q3) %>%
  summarise(
    count = n(),
    mean = mean(familiar, na.rm = TRUE),
    sd = sd(familiar, na.rm = TRUE)
  )
```

```{r familiarity-by-subfield-anova}
aov(familiar ~ Q3, data=survey_resp) %>% summary()
```

#### Summarize reproducible research practices ####
# there must be a prettier option for a crosstabulation with marginal sums!

summary(survey_resp$Q7a)
summary(survey_resp$Q7a_1)
ftable(data = survey_resp, Q7a_1 ~ Q7a)
summary(survey_resp$Q7a_3)
ftable(data = survey_resp, Q7a_3 ~ Q7a)
# FAMILIARITY with open source software outstrips USE of open source software
# and of documenting the computational environment
# low percentage of researchers "always" share information about computational environment

summary(survey_resp$Q7b)
summary(survey_resp$Q7b_1)
ftable(data = survey_resp, Q7b_1 ~ Q7b)
# FAMILIARITY with notebooks also outstrips their USE 
# less than 17% of researchers always use notebooks (lab, field, or computational)
# less than 40% use them most of the time

summary(survey_resp$Q7c)
summary(survey_resp$Q7c_1)
ftable(data = survey_resp, Q7c_1 ~ Q7c)
summary(survey_resp$Q7c_2)
ftable(data = survey_resp, Q7c_2 ~ Q7c)
summary(survey_resp$Q7c_3)
ftable(data = survey_resp, Q7c_3 ~ Q7c)
# most researchers are familiar with sharing/archiving data, but still only 16%
# report always doing so
# use of DOIs is inconsistent and use of metadata is rare

summary(survey_resp$Q7d)
summary(survey_resp$Q7d_1)
ftable(data = survey_resp, Q7d_1 ~ Q7d)
summary(survey_resp$Q7d_2)
ftable(data = survey_resp, Q7d_2 ~ Q7d)
# most researchers are at least somewhat familiar with sharing code or scripts
# but less than 40% actually do so all or most of the time
# and only 14% use version control software some or most of the time

summary(survey_resp$Q7e)
summary(survey_resp$Q7e_1)
ftable(data = survey_resp, Q7e_1 ~ Q7e)
# familiarity with pre-analysis plan registration is rare:
# 48% of respondents have no familiarity
# only 6 respondents reported using pre-analysis registrations all or most of the time

# In Sum, most geographers are at least somewhat familiar with open source software,
# data sharing, and code sharing.
# there are large gaps in implementation across all practices
# some practices are practically unknown in geography, including metadata, 
# version control software, and pre-analysis plan registration.

##### Summarize Beliefs #####

survey_resp %>% select(starts_with("Q8")) %>% summary()

# calculate familiarity with reproducible practices
survey_resp <- survey_resp %>% mutate(belief = as.numeric(Q8_1) + 
                                        (4 - as.numeric(Q8_2)) + 
                                        as.numeric(Q8_3) +
                                        as.numeric(Q8_4) + 
                                        as.numeric(Q8_5) +
                                        (4 - as.numeric(Q8_6))
                                      )
# a majority of geographers believe a failed reproduction does not disprove
# the original study, and geographers are split on whether it detracts from
# the original study's validity
# the most agreement is for students attempting reproductions as part of training

survey_resp %>% select(starts_with("Q8")) %>% lapply(as.numeric) %>% lapply(summary, na.rm=TRUE)
# we can see that in aggregate, geographers do not tend to strongly agree or disagree
# with beliefs about reproducibility
# most in favor of students conducting replications, and reproducibility contributing
# to credibility, and even some degree of compatibility with epistemology in the field
# however geographers are, on average, neutral about loss of validity for failed reproductions
# or loss of trust for studies without available data. 
# we disagree with a failed reproduction falsifying the original study.

# summarize and ANOVA beliefs with practices by subfield
group_by(survey_resp, Q3) %>%
  summarise(
    count = n(),
    mean = mean(belief, na.rm = TRUE),
    sd = sd(belief, na.rm = TRUE)
  )

aov(belief ~ Q3, data=survey_resp) %>% summary()
# again, significant differences in aggregate beliefs about the role of 
# reproducibility in subfields, with physical holding strongest affirmative
# beliefs, then methods, nature/society, and human geography