### 2023-05-29 docs/manuscript/manuscript.tex

\documentclass[]{interact}
\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage{subfigure}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

\usepackage{doi}
\usepackage{natbib}
\bibliographystyle{chicago}
\setcitestyle{authoryear,open={(},close={)}}
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,
    urlcolor=blue,
    citecolor=blue,
}

\usepackage{titlesec}
\titleformat*{\section}{\Large\bfseries}
\titleformat*{\subsection}{\large\bfseries}

\usepackage{endnotes}
\let\footnote=\endnote
\usepackage{etoolbox}
\patchcmd{\enoteformat}{1.8em}{0pt}{}{}

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

\usepackage{tabularx}
\usepackage{booktabs,caption}
\usepackage{threeparttable}

\usepackage{lscape}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\articletype{DRAFT MANUSCRIPT}

\title{Reproducible Research Practices and Barriers to Reproducible Research in Geography: Insights from a Survey}

\author{
\name{Peter Kedron\textsuperscript{a,b}\thanks{CONTACT Peter Kedron. Email: Peter.Kedron@asu.edu}, Joseph Holler\textsuperscript{c}, and Sarah Bardin\textsuperscript{a,b}}
\affil{\textsuperscript{a}School of Geographical Sciences and Urban Planning, Arizona State University, Tempe, Arizona, USA; \textsuperscript{b}Spatial Analysis Research Center (SPARC), Arizona State University, Tempe, Arizona, USA; \textsuperscript{c}Department of Geography, Middlebury College, Middlebury, Vermont, USA}
}

\maketitle

\begin{abstract}
While the number of reproduction and replication studies undertaken in the social and behavioral sciences continues to rise, such studies have not yet become commonplace in geography. 
Existing attempts to reproduce geographic research suggest that many studies cannot be fully reproduced, or are simply missing components needed to attempt a reproduction. 
Despite this suggestive evidence, we have not yet systematically assessed geographers' perceptions of reproducibility or the use of reproducible research practices across the discipline's diverse research traditions; nor have we identified the factors that keep geographers from conducting more reproduction studies.
We address each of these questions by surveying active geographic researchers selected using probability sampling techniques from a rigorously constructed sampling frame.
Our results identify a clear division in perception of reproducibility among geographic sub-fields. 
We also find varying levels of familiarity with reproducible research practices and a perceived lack of incentives to attempt and publish reproduction studies. 
Despite many barriers to reproducibility and divisions between subfields, we also find common foundations for examining and expanding reproducibility in the field.
These include interest in publishing transparent and reproducible methods, and in reproducing other researchers' studies for a variety of motivations including learning, checking the validity of a prior study, or extending the prior study.

\end{abstract}

\begin{keywords}
Reproducible Research, Epistemology, Geographic Research Methods
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section*{Introduction}
Reproducible research publicly discloses the evidence used to support claims made in prior work, facilitates the independent verification of those claims, and the extension of that work by the broader research community \citep{earp2015, nosek2012scientific, Schmidt2009}.
While reproducibility is not a guarantee of scientific or practical usefulness, it does provide a strong basis for the collective evaluation of ideas.
Nonetheless, systematic reviews \citep{moraila2014measuring, iqbal2016reproducible} of published research papers reveal a lack of data and code for reproducibility  and attempts to reproduce published research papers \citep{raghupathi2022reproducibility, chang2015economics} frequently fail.
Previous studies consistently link the irreproducibility of research to inadequate record keeping, opaque reporting, the inaccessibility of research components, and a lack of incentives to share research details or to attempt reproduction studies \citep{ranstam2000fraud, anderson2007normative, NASEM2019}.
Surveys of researchers find that few researchers are attempting to independently reproduce the work of others \citep{baker20161, boulbes2018survey}.
At the same time, a sizable portion of researchers admit engaging in questionable research practices tied to the publication of false positive results and low reproducibility rates \citep{fanelli2009many, fraser2018questionable}.

Despite these concerns, the available literature currently provides insufficient evidence to conclusively evaluate the reproducibility of research generally, or disciplinary research specifically.
This knowledge gap exists because researchers conducting reproducibility assessments, surveys of researcher practices, and reproduction attempts have not defined clear target populations for inference, or used probability sampling techniques structured by well-defined sampling frames that support generalization \citep{NASEM2019}.
Attempts to reproduce past results typically focus on recreating the results of one or a small number of studies selected based on topical interest or researcher familiarity \citep{camerer2016evaluating, camerer2018evaluating, open2015estimating}. 
Attempts to catalog or survey reproducible research practices have drawn samples of research papers from archives such as conference paper series, specific journals, or disciplinary repositories \citep{byrne_2017, gundersen2018state, stodden2016enhancing, stodden2018enabling}, or have drawn samples of researchers from specific journals, members of professional associations, or conference attendees \citep{baker20161}.
These samples are convenient, but are not guaranteed to be representative of researchers and research activity in a particular field. 
Surveys have also commonly failed to systematically report the methodological details (e.g., response rate) needed to assess and address potential bias in survey response, which further complicates their evaluation and use. 
Finally, both reproduction attempts and reproducibility surveys have focused on evaluating the computational components of past studies such as data and code availability rather than all aspects of research design and execution.

Reproducibility surveys and reproduction attempts in the geographic literature face these same challenges. 
The few available reproducibility surveys in the discipline have relied on convenience samples drawn from specialist conferences and have only focused on computationally intensive forms of geographic research \citep{balz2020reproducibility, konkol2019, ostermann2017}. 
The small number of published attempts to reproduction geographic research similarly have focused on the computational reproducibility of conference papers \citep{Nust-AGILE_2018, Nust_AGILE_2020, Nust_AGILE_2021, Nust_AGILE_2022, ostermann2021,} or on specific topics such as COVID-19 \citep{Holler2023disability, Kedron2022dimaggio, Kedron2023Beyond, paez2022reproducibility}. 
More recent reproduction attempts by \citet{Kedron2023Beyond} show that the factors hindering the recreation of results and the evaluation of claims likely extend beyond computation into the conceptualization and design of geographic research.

Geographers continue to debate the role of reproduction studies in the discipline \citep{brunsdon2016, goodchild2021Annals, kedron2022replication, kedron2021IJGIS, singleton2016, sui2021, Wainwright2021}, examine the reproducibility of individual studies \citep{Nust_AGILE_2022, ostermann2021}, and build the infrastructure needed to support reproducible research \citep{nust2019, wilson2021, yin2019cybergis}. 
And yet, we still have not systematically assessed the use of reproducible research practices across the discipline's diverse research traditions, or identified the factors that have hindered geographers from adopting reproducible research practices and conducting reproduction studies. 
Without answers to these questions it is unclear what actions geographers should take if they wish to improve the reproducibility of work in the discipline.

To address this gap in our collective knowledge, we surveyed geographic researchers about their understanding of reproducibility, perception of reproducibility in their sub-fields,  familiarity and use of reproducible research practices, and the factors they see as barriers to reproducible research.
To support generalization, we used a probability sampling scheme designed to capture the breadth of geographic research.
In the remainder of this paper, we first present the design of our survey, sampling strategy, and analytical approach. 
We then present the results of our survey, first focusing on researcher perceptions and use of reproducible research practices, and then analyzing researcher experiences attempting to reproduce prior work. 
Finally, we discuss the implications of our survey results and conclude by presenting a snapshot of geographic researcher perceptions of reproducibility and where the discipline may go from here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Data and Methods}
Complete documentation of the procedures, survey instrument, and other materials used in this study are available through the Survey of Reproducibility in Geographic Research Repository (\citet{Kedron_Holler_Bardin_Hilgendorf_2022} - \url{https://osf.io/5yeq8/}) hosted by the Open Science Framework (OSF).
The OSF repository connects to a GitHub repository which hosts the anonymized dataset and code used to create all results and supplemental materials along with a complete history of their development. 
All of the results presented in this paper can be independently reproduced using the materials in that repository.
We encourage interested readers to critically evaluate and build on these materials through their own work.
Before the start of data collection, we registered a preanalysis plan for the survey with OSF Registries (\citet{Kedron_Survey_PAP} - \url{https://osf.io/6zjcp}). 
The survey was conducted under the approval and supervision of the Arizona State Institutional Review Board - \textit{STUDY00014232}.

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Sampling Frame}
Our target population of interest is researchers who have recently published in the field of geography. 
We followed a 4-step procedure to create a sampling frame for our survey that captures this diverse population of researchers and the approaches they use when studying geography.

First, beginning at the publication level, we identified journals indexed as either geography or physical geography by the \href{https://access.clarivate.com/}{Web of Science's Journal Citation Reports} that also had a 5-year impact factor greater than 1.5.
From those journals, we created a database of all articles published between 2017 and 2021.

Second, we used Arizona State University's institutional subscription to the \href{https://www.scopus.com/home.uri}{Scopus Database} to extract journal information (e.g., subject area, ranking), article information (e.g., abstract, citation counts), and author information (e.g., corresponding status, email) for each publication. 
Because our intention was to capture individuals actively publishing new geographic research, we retained publications indexed by Scopus as \textit{document type = ``Article''} and removed all other publication types (e.g., editorials, book reviews) from our article database. 
We also removed articles with missing authorship information.

Third, we created a list of researchers and their published articles, focusing on corresponding authors for two reasons.
(1) Corresponding authorship is one indicator of the level of involvement an individual had in a given work. 
While imperfect, it was the best available indicator in the Scopus database as across journals there is no commonly adopted policy for declarations of author work (e.g., CRediT Statements).
(2) Scopus maintains email contact information for all corresponding authors, which gave us a means of contacting researchers in our sampling frame.
Scopus also maintains a unique identifier for each author (author-id) across time, which allowed us to identify authors across publications.

Fourth, we constructed a sampling frame of unique researchers and their most recent email contact information. 
We determined uniqueness by grouping researchers by their author-id, and we determined the most recent contact information by selecting records associated with the most recent year of publication. 
For 383 researchers who had two or more distinct emails in the latest year of publication, we removed non-institutional personal email addresses and then selected one of the remaining institutional email address.

Applying these criteria yielded a sampling frame of 29,828 researchers. 
On average, these authors published 2.7 articles in geography journals meeting our criteria between 2017 and 2021. 
Roughly one-third (33.0\%) were most recently a corresponding author for an article published in a general geography journal. 
A similar proportion (32.0\%) were most recently a corresponding author for an article published in an earth sciences journal, and smaller proportions had published in the social sciences and cultural geography (20.0\% and 16.0\%, respectively).

\subsection*{Survey Instrument}
The survey first established eligibility based on age and geographic research activity in the past five years and asked researchers to report their primary subfield and methodology.
We asked each participant to assess their familiarity with the term ``reproducibility'' and to provide their own definition. 
We then provided a definition based on the \citet{NASEM2019} to establish a common understanding of reproducibility for the remainder of the survey.
Remaining questions assessed familiarity and use of reproducible research practices (22 questions), perceptions of the reproducibility of geographic research (2 questions), and beliefs about reproducibility with regards to its significance (17 questions) and barriers (13 questions).
For researchers who reported attempting reproductions, we asked researcher to elaborate on their motivations and outcomes (9 questions).

We developed the survey questions following a review of prior reproducibility surveys \citep[e.g.,][]{fanelli2009many,baker20161, konkol2019} and our own reading of recurring issues in the reproducibility literature. 
We pilot tested the survey instrument with \textit{n}=19 graduate students and geography faculty with differing levels of experience, topical focus, and methodological background. 
After pilot testing, we removed these individuals from our sampling frame to ensure they would not be included in our final sample.

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Data Collection}
We used a digital form of the Tailored Design Method \citep{dillman2014internet} to survey geographic researchers between May 17 and June 10, 2022.
A simple random sample of 2,000 researchers was drawn without replacement from our sampling frame, and those researchers were invited via email to participate in the online survey. 
Researchers received their initial invitation on May 17, 2022. 
Two reminder emails were sent to researchers that had not yet completed the survey on May 26 and May 31, 2022.

The online survey was administered through \href{https://www.qualtrics.com/}{Qualtrics}. 
Participation in the survey was entirely voluntary. 
Each researcher that opted to participate in the survey was provided with IRB approved consent documentation and linked to the internet survey instrument. 
Participants were also given the option to provide an email address for eligibility for one of three  prizes of 90 US\$, selected randomly after the data collection period.
Participating researchers had the option to exit and re-enter the survey and were also able to review and change their answers using a back button as they progressed through the survey.
At the end of the data collection period, responses were checked for completeness and coded using the reporting standards of the American Association For Public Opinion Research \citep{aaporstandards}.
Responses were downloaded from Qualtrics, anonymized, and stored in a public de-identified database in the research compendium (self-identifying reference removed).

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Analytical Approach}

We conducted two statistical analyses of the survey responses.
First, we analyzed researcher perspectives on reproducibility by coding and calculating summary variables and statistics for three themes: how geographic researchers define reproducibility, familiarity and experience with reproducible research practices, and perceived barriers to reproducibility.
Second, we analyzed researchers' experiences with reproducing prior studies, for those who had done so, including their motivations and experience of successes and barriers.

\subsubsection*{Analyzing Researcher Perspectives on Reproducibility}

\noindent For our first set of analyses, we examined the full set of survey responses.

\textit{Defining Reproducibility:} 
We coded participants' qualitative definitions of ``reproducibility'' following two procedures.
First, we measured the similarity of each provided definition to the definition adopted by the \citet{NASEM2019}. 
The NASEM defines reproducible research as having four characteristics --- same data, same procedure, same results, and same conditions.
To make this comparison, each of the authors independently coded each respondent definition for the presence/absence of each of the four characteristics included in the NASEM definition.
Disagreements in the assignment of codes were resolved through discussion between the three authors.
We created an aggregate measure of definition similarity for the final coded response for each participant by counting the presence of each NASEM definition characteristic, resulting in a measure with the domain [0, 4].

Second, we coded each definition to one of four motivations for ensuring the reproducibility of a study: (i) to facilitate the assessment of prior work, (ii) assess experimental research, (iii) to improve transparency and facilitate further extension of work, and (iv) to improve the transparency and consistency of data collection.
We derived this coding from common themes in the responses and our own reading of the reproducibility literature.
As above, each definition was independently coded by each author before code assignments across authors were compared with disagreements resolved through discussion.

\textit{Familiarity and Experience:} We measured participant familiarity and experience with five reproducibility enhancing research practices: the adoption of open-source software, research notebooks, data sharing, code sharing, and research plan preregistration. 
We created a measure of familiarity by counting the variety of practices a researcher was `somewhat' or `very' familiar with.
We measured experience by counting the variety of practices a researcher reported using `most of the time' or `always' when conducting their own research. 
Both measures have domains [0, 5].

\textit{Barriers:} Finally, we asked asked respondents to identify how frequently they believe 12 different factors contributed to a lack of reproducibility in the discipline. 
From those responses, we created a measure of perceived barriers by counting the number of factors a researcher believes `frequently' or `occasionally' contributed to irreproducibility, resulting in a measure with domain [0, 12].

To examine variation in researcher understanding of reproducibility across disciplinary subfields and methodological traditions, we cross tabulated these summary measures and responses to individual questions for these subgroups. 
We then compared response frequencies across subgroups.

%%%%%%%
\subsubsection*{Analyzing Researcher Reproduction Attempts}
\noindent For our second set of analyses, we examined only the responses of researchers who reported attempting a reproduction in the past two years to understand what motivated reproduction attempts, how successful those attempts were, and what factors hindered success.

\textit{Motivations:} To assess what motivated researchers to attempt reproductions, two of the authors independently assigned participants text-based response to groups with the following motivations: i) verify/check published research, ii) learn from published research for extension or teaching, iii) internally check their own research to verify their work and/or increase the transparency of their work, and iv) replicate a study with new data.
We identified disagreements in motivation assignments, and the third author again resolved all disputes in consultation with the two assigning authors.
We selected groups i) and ii) for further analysis of reproduction attempts, as these motivations most closely matched our definition of reproduction.

\textit{Success and Barriers:} To analyze participant success, we create statistical summaries for a series of questions that asked researcher to identify whether they were able to partially or completely recreate some or all results of the target study. 
We similarly analyzed barriers to the reproduction of results by creating statistical summaries of participants' ability to access key study artifacts (e.g., data, procedural information, code).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
A total of \textit{n}=218 of the authors we contacted completed the online survey with information sufficient for analysis. 
The contact rate for the survey was 13.9 percent and the cooperation rate was 77.6 percent, yielding an overall response rate of 10.8 percent. 
The refusal rate was 3.1 percent\endnote{All outcome rates are reported using \citet{aaporstandards} standards. 
The outcome rates used were - response rate 2, cooperation rate 2, refusal rate 1, and contact rate 1.}.
Another \textit{n}=40 authors started the survey, but did not complete enough of the survey to be included in the analysis.

Respondents were predominantly male (65.1\%) and between the ages of 35 and 55 (62.4\%). 
The majority of respondents were academics, and they were balanced across career levels from graduate students to full professors with no one career level comprising more that 30 percent of the sample.
Respondents identified with each of the four major disciplinary subfields---physical geography (29.8\%), geographic methods and GIScience (28.0\%), nature and society (10.1\%), and human geography (30.7\%); and three major methodological approaches---quantitative (42.2\%), mixed methods (39.0\%), and qualitative (18.3\%).

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Researcher Perspectives on Reproducibility}
Reproducibility is on the minds of geographic researchers.
Nearly all researchers reported being at least somewhat familiar with the term reproducibility (89.0\%), with half reporting being very familiar with the term (53.6\%).
More pointedly, the majority of survey respondents reported thinking about the reproducibility of their own research (80.7\%), discussing reproducibility with a colleague (70.6\%), and questioning the reproducibility of published work (57.3\%) in the past two years. 
More than half of the researchers we surveyed (52.8\%) also reported considering reproducibility while peer reviewing a grant proposal or publication during the same time frame. 
However, researchers estimated that only 50.6 percent of the results published in the discipline were reproducible, albeit with a large standard deviation of 24.7 percent that suggests a great deal of uncertainty about the true value. 
Few respondents reported attempting to reproduce the work of other researchers (14.7\%) with fewer still attempting to publish those reproduction studies (6.8\%). 
Strikingly, only a small majority of respondents believe reproducibility is compatible with epistemologies used in their disciplinary subfields (58.2\%).

Table \ref{tab:overview} summarizes how researchers define reproducibility, their familiarity with reproducible research practices and experience using them, and the factors they see as barriers to reproducibility in geography. 
We present the mean and standard deviation for the overall sample and for each subfield and methodological approach.
In aggregate, the data reveal trends in definition, familiarity, experience, and barriers of reproducibility between the sub-disciplines and methodological approaches. 
Respondents that self-identified as specializing in physical geography and geographic methods consistently report higher indicators than those working in nature and society and human geography.
Similarly, respondents that identified as primarily using quantitative and mixed methods approaches consistently report higher indicators than those using qualitative methods.
The following sub-sections present detailed results for each of these topics, highlighting the principal sources of difference between sub-disciplines and methodological approaches.

\begin{center}
\textbf{Insert Table \ref{tab:overview} About Here}
\end{center}

%%%%%%%%%%%
\subsubsection*{Definitions and Importance of Reproducibility}
A total of 181 (83.0\%) of our survey respondents provided an interpretable definition of reproducibility.  
Geographic researchers provided definitions of reproducibility that explicitly included an average of 1.83 of the four characteristics from the definition adopted by the NASEM.
The availability and use of the same research procedures (80.7\%) and results (74.0\%) were the characteristics of reproducibility most frequently identified by researchers. 
Less than half of respondents explicitly included use of the same data (38.1\%) or the need to work in the same context (17.7\%) in their definitions. 
The pattern of similarity to the NASEM definition and each of its components was consistent across subfields and methodological approaches, with a slightly greater emphasis on data and procedural availability amongst quantitative and methods-focused researchers.
The lower inclusion of data and context in definitions may be explained by researchers conceptualizing reproducibility as the formal NASEM definition of replicability, which emphasizes the testing of similar questions and procedures in new contexts with new data. 
For example, one respondent defines reproducibility as, ``the extent to which the research design can be replicated in different geographical contexts."
We observed this alternative definition of reproducibility in 20.4\% of respondents' definitions.

Researchers' definitions of reproducibility were primarily connected to two epistemic functions.
Just over half of respondents (52.5\%) defined reproducibility as a means of assessing prior work for errors or inconsistencies through comparison of original results to results from an attempted reproduction.
These ranged from rigid bitwise quantitative interpretations, as in the ``ability to re-generate exactly the results published based on the data and code provided by the authors", to more flexible interpretations in which ``other researchers could use the same or similar methodology without great difficulty and, given similar data, arrive at comparable results."
Responses included definitions with a focus on experimental science, as in ``an ability to produce consistent results when an experiment is repeated!"

Nearly all other researchers (40.9\%) tied reproducibility to the need for transparency in research so that others could independently expand upon past studies.
For example a quantitative geographer stated: ``The methods should provide sufficient information to be able to reproduce the results. In quantitative science this should, at minimum, provide all the equations and algorithms used for any calculation. In the interest of increasing transparency in science, the practice of sharing the code should be encouraged."
For others, open science did not necessarily need to result in identical results, reflected in this definition: ``As a qualitative researcher doing in-depth case study research, my studies can not be perfectly reproduced. But reproducibility sits in the openness about methods and data collection practices, as well as critical reflection about strengths and weaknesses of my research. When we write about those things in the methods section in our papers and theses, their reproducibility is increased."

The remainder of responses (6.6\%) emphasized repeatable or reliable data observation over all other dimensions of reproducibility.
For example, ``As a historical geographer, working with qualitative research methods, I understand reproducibility more in terms of sources than of methods. I see reproducible research as being that which makes clear the origin and location of its data."
A physical geographer similarly emphasized data observations: ``Data/observations of some variable can be recovered repeatedly by different observers/methods."

Responses to related Likert questions from the full survey sample (\textit{n}=218) support the results from the subsample of 181 qualitative definitions analyzed above.
A majority of researchers identified reproducibility as important for validating (75.2\%) and establishing the credibility (72.5\%) of research.  
Respondents also saw reproducing studies as important to reducing the presence of persistent errors in the discipline (77.5\%) and to increasing trust in research findings (78.5\%).
In parallel with the need for openness and transparency in science, most respondents agreed with the importance of reproducibility for research efficiency (63.3\%), communication with academics (68.8\%) and practitioners (64.7\%), and training students (75.7\%).

Despite wide recognition of reproducibility as epistemically important, respondents were cautious about drawing conclusions from a single study or reproduction attempt. 
Only half of the respondents (50.9\%) agreed that when researchers don't share their data they have less trust in a study.
A smaller percentage (41.7\%) agreed that inability to reproduce a result detracts from the validity of a study, and an even smaller minority agreed that such inability implies that the result is false (26.2\%).

Qualitative researchers identified reproducibility as playing a much smaller epistemic role compared to the discipline as a whole. 
A majority of the qualitative researchers responded that reproducibility is incompatible with the epistemologies used in their subfield (75.0\%).
Accordingly, a small percentage of qualitative researchers agreed that reproducibility is important for validating research (25.0\%) or establishing its credibility (20.0\%).
Qualitative geographers similarly placed less emphasis on reproducibility as a means of increasing the accessibility and extensibility of research.
Few qualitative researchers had less trust in a study when researchers did not share their data (27.5\%), or saw reproducibility as important for sharing research with academics (27.5\%).

The data from our sample of researchers has shown broad recognition of reproducibility and its importance in geography with three caveats---confusion between reproducibility and replicability, different perspectives from researchers using qualitative methods, and caution about judging the trustworthiness or validity published research based on its reproducibility. 
In this context, are individual researchers aware of the research practices needed to enhance reproducibility, and have these practices already been adopted for use in research?

%%%%%%%%%%%
\subsubsection*{Familiarity and Experience with Reproducible Research Practices}
Geographic researchers were familiar with an average of 3.26 different reproducible research practices, but only reported experience using an average of 1.44 of these practices in their own work.
More than half of all researchers reported familiarity with with data sharing (86.7\%), open source software (85.3\%), field and lab notebooks (67.0\%), and code sharing (59.2\%).
However, a far smaller number of researchers reported using these ``familiar" practices regularly in their own work. 
Less than half of the researchers surveyed reported sharing their data (44.5\%), using open source software (38.1\%), using field or lab notebooks to record their work (40.0\%), or sharing their code (18.8\%) most or all of the time. 
Only a small subset of researchers reported familiarity with the pre-registration of research designs and protocols (27.5\%), or regular use of this practice (2.7\%).

Researcher familiarity and use of reproducible research practices varied by disciplinary subfield and methodological approach. 
Researchers that identified as physical geographers or methodologists and GIScientists reported being familiar with one to two more reproducible researcher practices than human geographers and those focused on nature and society.
Researcher practices similarly diverged by subfield, but no subset of researchers reported using on average more than two of these practices regularly in their work.  
When grouped by methodological approach, quantitative and mixed-methods researchers on average reported familiarity with and use of two more reproducible research practices when compared to qualitative researchers.

Differences in researcher familiarity and use of specific reproducible research practices across subfields and approaches was greatest for practices more typical of quantitative workflows. 
For example, 96.7 percent of quantitative and 89.5 percent of mixed-methods researchers reported familiarity with open source software, while 52.5 percent of qualitative researchers with the same resources.
These differences were greater for code sharing practices. 
Just 12.5 percent of qualitative researchers reported familiarity with code sharing, while 81.5 percent of quantitative and 55.7 percent of mixed-methods researchers reported familiarity with the same practice.
However, even among quantitiative and mixed-methods researchers, familiarity with reproducible research practices did not translate into regular use of those practices.
Just over half of quantitative (51.2\%) and mixed-methods (51.8\%) researchers reported sharing their data most of the time.
Less than a third of these same researchers reported sharing their code most of the time. 
These rates were lower for qualitative geographers, of which only a small subset reported using open source software (10.0\%) or field notes (20.0\%), or sharing data (12.5\%) and code (5.0\%) most of the time.

\subsubsection*{Perceived Barriers to Reproducible Research}
Aligning with the differences we observed between researcher familiarity with reproducibility enhancing research practices and researcher use of those same practices, respondents to our survey identified an average of 8.20 factors contributing to a lack or reproducible research across the discipline. 
Geographic researchers working in different subfields and with different methodological approaches each identified a similar number of barriers to reproducibility (Table \ref{tab:barriers}).
Qualitative geographers may be the one group that deviates from this consistent pattern. 
These researchers identified fewer barriers to reproducibility on average, but with a greater variance that left us unable to distinguish this group from any other.
To examine differences in researcher beliefs about specific factors potentially hindering reproducibility in the discipline, we divided the 12 factors we examined into three groups---those related to the research environment, the availability of research artifacts, and study-specific characteristics.

\begin{center}
\textbf{Insert Table \ref{tab:barriers} About Here}
\end{center}

Geographic researchers identified the incentive structure of the researcher environment as an important barrier to reproducibility in the discipline.
A majority of geographic researchers identified both the pressure to publish original research (71.5\%) and insufficient oversight of the research process (71.1\%) as barriers to reproducibility.
A high percentage of physical, methods-focused, quantitative, and mixed-methods researchers identified these factors as barriers. 
Qualitative geographers were least concerned with these factors, but a majority of these researchers still identified both factors as barriers to reproducibility in the field.
A minority of geographic researchers (28.4\%) believe that the fabrication of data, the manipulation of research results, and similar forms of fraud are a cause of irreproducibility in the discipline.
This percentage is consistent with large surveys and meta-analyses of research on scientific fraud across other scientific disciplines \citep{fanelli2009many, baker20161}.

Researchers identified the unavailability of research artifacts (e.g., data, code) as a second barrier to reproducibility, but the importance placed on different artifacts varied by subfield and methodological approach.
A higher percentage physical and methods-focused researchers identified all five of the artifacts we investigated as common barriers to reproducibility as compared to human and nature-society researchers.
The largest differences between these groups existed in researchers beliefs about how often the availability of research protocols/code and the use of restricted data or software impacted reproducibility.
More than three-quarters of physical (77.0\%) and methods-focused (86.9\%) geographers identified the availability of protocols/code as contributing to irreproducibile results, whereas less than two-thirds of human (59.7\%) and nature-society (50.0\%) identified this same factor. 
Only a large percentage of methods-focused researchers (82.0\%) identified the use of restricted data/software as a common contributor to irreproducibility in the discipline.
Intuitively, a similar gap existed between qualitative researchers and mixed-methods or quantitative researchers with regards to identifying code availability or the use of restricted dara/software as contributing to irreproducibility.

A majority of researchers identified geographic variation (71.5\%), researcher positionality (64.2\%), and chance (62.3\%) as study-specific factors limiting the reproducibility of geographic research.
Minor variations in the emphasis placed on these factors exist across subfields and approaches. 
A higher percentage of nature-society (81.8\%) and physical (80.0\%) researchers emphasized the important role that spatial variation and complexity of geographic processes can play when attempting to reproduce geographic research, but this factor was also recognized by researchers across subfields and approaches. 
A smaller percentage of physical geographers placed emphasis on the impact of researcher positionality may have on reproducibility (50.8\%) when compared to all other subfields.  
Differences between the computational environment used to conduct an original study and a reproduction attempt was generally not seen as a factor contributing to a lack of reproducibility in the discipline. 
Of all subgroups, only a majority of methods-focused (68.9\%) and quantitative researchers (63.1\%) were concerned with computational environments, reflecting research practices used in their areas of research.

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Attempted Reproductions}
A total of 102 of the researchers that responded to our survey (46.8\%) reported attempting a reproduction study during the past two years.
However, 23 of those researchers were reproducing their own research results, while another 13 were replicating prior studies in new locations.
In the end, only \textit{n}=32 (14.7\%) of all respondents reported attempting to reproduce a study originally conducted by another researcher during the past two years.

This subset of 32 participants formed the basis for our second analysis of researcher practices and experiences when attempting reproductions of the work of others.
Reproduction attempts were predominantly made by geographic researchers that self-identified with the physical geography (43.8\%) or geographic methods (37.5\%) subfields.
Respondents attempting reproductions were also focused on quantitative (68.8\%) and mixed-methods (41.2\%) approaches. 
Only eight of the researchers that attempted reproductions reported submitting any of their findings for publication.

Most of the 32 researchers that attempted to reproduce a prior study reported at least some success in accessing data and procedures and in reproducing the prior study results.   
The majority of researchers (87.5\%) were able to access some of the data used in the original study, but few researchers (12.5\%) reported access to all of the original data.
Researchers also reported the ability to access at least some information about the study procedures (68.8\%) and computational environment (59.4\%), but limited ability to access all procedural (9.4\%) and computational environment information (12.5\%).

Reproduction attempts may produce results for comparison to \textit{some} or \textit{all} of the results in a prior study.
A reproduction may be \textit{identical} by finding the exact same results, or may be \textit{partial} by finding slightly different results that still support the same conclusions.
Nearly all researchers reported at least partially reproducing some results (81.3\%), but only seven (21.9\%) reported being able to at least partially reproduce all results.
Only three researchers (9.4\%) were able to identically reproduce all results.

Access to prior study data and procedural information appears to impact the ability to reproduce prior study results. 
When researchers had access to some of the data from the original study, they reported being able to at least partially reproduce all results in 6 of 24 instances. 
That success rate rose to 3 of 4 when researchers reported access to all data. 
We observed a similar increase in researcher ability to at least partially reproduce all results when researchers reported access to all procedural information. 
When researchers had access to some of the procedural information from the original study, they reported being able to at least partially reproduce all results in 6 of 19 instances.
That success rate rose to 3 of 3 when researchers reported access to all procedural information and all code.
The small number of reproduction study attempts reported in the survey results make it difficult to draw broad conclusions.
However, the results are internally consistent and intuitively support the importance of available data and procedures for successful reproductions of geographic research.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion and Limitations}
Our survey results indicate that geographic researchers are aware of reproducibility and reproducible research practices, but have yet to incorporate many of those practices into their own work. 
We found that few researchers attempt to independently reproduce the work of others, or to publish the reproduction attempts they do undertake.
In alignment with the broader reproducibility literature, geographic researchers identify the lack methodological transparency and the unavailability of data and procedural information---particularly code---as key barriers to reproducibility in the discipline. 
Our results also suggest the existence of an incentive problem within the discipline, as pressure to publish original research and a lack of oversight create the opportunity to publish irreproducibile work and limited reward for attempting and publishing reproduction studies.  
These findings empirically confirm prior speculations about the reproducibility of geographic research---on the whole, geographers have yet to make reproducible research practices and reproduction attempts a regular part of disciplinary practice.

Our findings also suggest that geographic researchers do not share a single definition of the reproducibility. 
While researchers share beliefs about the epistemological functions of independent reproductions, they provide definitions that contained different requirements for similarity across studies in terms of data, procedures, results, and context.  
Moreover, a subset of researchers define reproducibility as what the \citet{NASEM2019} defines as replicability---the ability to obtain consistent results across studies designed to answer the same question, each of which has obtained its own data.
The interchangeable use of reproducibility and replicability, or the outright reversal of definitions we observed in our sample, has also been documented across the sciences \citep{barba2018terminologies, plesser2018reproducibility}. 
Given that geography has no established standard use of either term and that many geographic researchers are also trained in other disciplines, it is likely that researchers at least partially inform their definition of reproducibility using concepts prominent in their cognate fields.

More broadly, the observed variation in terminology we observed is important for at least two reasons. 
First, variation in geographic researchers' understanding of reproducibility reflects the discipline's diverse traditions and ways of knowing. 
Acknowledging this diversity as a strength of the discipline, productive discussions about reproducibility should consider how reproducible research practice fits into different traditions and what common understanding exists across traditions. 
Second, if researchers lock into an ontological debate about the definition of reproducibility, the community may hinder a more productive discussion about the epistemological role independent reproductions and open science practices can or should play in the discipline.

Our findings point to potentially productive pathways for such a discussion. 
Even among the subset of respondents most uncertain about the role of reproducibility in geographic research---qualitative geographers---we observed recognition of the value reproducible research practices bring to the transparency of research, with a particular emphasis on the sharing of methods and procedures.   
Even acknowledging that spatial heterogeneity and the uniqueness of places can complicate how reproduction and replication studies can be used to build geographic knowledge, conversations to address that challenge likely best begins with some common understanding of what role they are meant to play in that process.

Our results suggest that qualitative researchers have a different perspective on the role or reproducibility in geographic research when compared to other subgroups.
The extent to which we can generalize this perspective is limited by the size of our sample of qualitative researchers (\textit{n}=40).  
Nonetheless, the majority of our qualitatively-focused respondents believe reproducibility is incompatible with their epistemological approach.  
The reproducibility definitions these researchers provided suggest this belief may be rooted in the role researcher positionality plays in the collection and interpretation of qualitative data. 
Several respondents questioned whether an independent researcher analyzing the same data through their unique interpretive could come to the exact same results. 
Respondents also focused on what one researcher termed 'transferability', the desire to know that what they have discovered is relevant in other similar contexts.

Engaging qualitative researchers and integrating their perspectives and knowledge has the potential address issues in the reproducibility literature.  
Perhaps one way forward is to initiate a conversation that highlights how reproducibility is not an absolute standard or determinant of research quality, but instead a means of clarifying for others what was was done in a study and why conclusions were drawn as they were.
Even if a researcher believes positionality influences data collection and interpretation, using reproducible research practices to control all variables of research design \textit{except} researcher positionality may help convey that position and its impact on study results.
To our knowledge, there is little explicit discussion in the reproducibility literature about how researcher positionality can or should be recorded and conveyed to other researchers. 
It would also be interesting to know whether certain elements of researcher positions appear stable across researchers and contexts.
Such work could also move the conversation about reproducibility away from a current focus on exactly recreating numerical results, and back to the practice's deeper function---independently assessing the claims of prior research.

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Limitations}
To our knowledge, our work is the first systematic attempt to survey geographic researchers about reproducibility. 
To draw a reliable and generalizable understanding of this issue, we developed a sampling frame using recent publications in indexed journals to gather information from our target population of individuals actively publishing geographic research. 
Ideally, we would stratify this set of potential respondents into subgroups based on common characteristics meaningful to their knowledge of reproducibility and use of reproducibile research practices, and then randomly draw participants from these subgroups to create a balanced and representative sample. 
If our resulting sample was imbalanced across subgroups, we would then use a post-stratification procedure to adjust the non-representative sample using the predictors known to characterize the strata and response.

We could not follow this approach for three reasons. 
First, meaningful stratification and post-stratification require knowledge of what predicts differences in response.
Given the currently limited understanding of reproducibility within geography, we can only speculate about the researcher characteristics predictive of different levels of familiarity and experience with reproducible research, e.g. subfield, methodological approach, or career position.
We do not have the foundational knowledge needed to identify reliable predictors.
Inverting the issue our survey as the first discilpine-wide measurement of these issues contributes to the development of this foundational knowledge to support future studies.
Second, even if predictors could be selected to be used in stratification or post-stratification, they must be measured for the population of geographic researchers.
We are not aware of any census of geographic researchers that could be used to identify and classify all those working in the field.
Third, given population-wide measures of key predictors, it would be necessary to have measures of those predictors for potential respondents, which can be difficult within geography.
For example, divisions between subfields and approaches in geography are often unclear and researchers often identify with multiple subfields and approaches.

Given the limitations to stratifying or balancing a survey on reproducibility in geography, our study should be viewed as an exploratory analysis with random sampling and a transparent, reproducible methodology for sample frame construction.

Absent stratification, we have taken steps to reduce exclusion bias by including in our sampling frame all researchers publishing as corresponding authors in any one of X geography journals over a five year period.
While we cannot eliminate the possibility of self-selection bias from our survey, we attempted to quantify potential self-selection by examining the characteristics of researchers who started our survey but did not complete the full survey. 
For 29 researchers who reported their subfield but did not complete the full survey, completion rates for all subfields were between 84 and 87 percent, except slightly higher rates for methods / GIScience (96.8\%).
Considering 31 researchers who reported their method but did not complete the full survey, completion rates were 84.2 percent for mixed methods, 87.0 percent for qualitative methods, and 91.1 percent for quantitative methods. 
These values suggest that self-selection was not a significant issue.
Finally, we attempted to mitigate the potential for questionnaire bias, which could be caused by partially basing our survey instrument on prior studies that over-represent perspectives from the computational and experimental sciences.
To address this concern, we incorporated into our survey questions from a parallel review  of the reproducibility literature available within geography and a review of critiques of positivist science made by social scientists and human geographers, and included space for qualitative responses in each survey theme, and pilot tested out instrument with a diverse set of geographers.

%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Conclusion}
Reproducibility remains an emerging topic in the geographic literature. 
In this study, we provide the first systematic survey of the use of reproducible research practices across geography's diverse research traditions. 
Our results make clear that geographic researchers are aware of reproducible research practices, but lack direct experience using those practices.
Academic incentive systems and the inaccessibility of key components of prior research hinder reproducibility, and a small percentage of researchers are attempting to independently reproduce past work.

Arising from survey results, we see two major opportunities for geographers to contribute to the interdisciplinary challenges and debates surrounding reproducibility.
Geographers may use the methodological and epistemological diversity in our discipline to pragmatically de-emphasize target fixation on computational reproducibility and refocus debates about reproducibility on practical ways in which attempting to reproduce the findings of prior studies may enhance the accumulation of knowledge and learning across different epistemologies.
By extension, such a pragmatic debate would naturally extend back to approaches to designing and communicating geographic research to facilitate meaningful reproducibility across methodologies and disciplines and beyond the limited scope of computational reproducibility.
Across subfields and methods, geographers are aware of the challenges that spatial heterogeneity and uniqueness of place may pose for reproducibility, and thus we are well-positioned to infuse the theory and practice of reproducibility with geographic context.
Both of these contributions are particularly salient in the context of convergence research addressing the most urgent challenges facing humanity, including global health, climate change, poverty, conflict, and resilience and transformation in rapidly changing environments.

Our work also creates a foundation for the further empirical investigation of reproducibility within geography and across the discipline's many traditions. 
We have made all the materials used in the development and execution of this research openly available so others can critique and extend our work.
We urge other researchers to replicate our research, improve our sampling frame and survey instrument, and progressively create a deeper understanding of questions we only begin to address in this work.

%%%%%%%%%%%%%%%%%%%%%%%%
\theendnotes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgement(s)}
We thank Tyler Hoffman for providing technical assistance in the development and execution of a set of trial queries using the Scopus API.

\section*{Funding}
This material is based on work supported by the National Science Foundation under Grant No. \textbf{BCS-2049837}.

\section*{Notes on contributor(s)}
\textbf{Kedron:} Conceptualization, Methodology, Writing - Original Draft, Writing - Review and Editing, Supervision, Project Administration, Funding Acquisition. \textbf{Holler:} Conceptualization, Methodology, Data Curation, Writing - Review and Editing, Funding Acquisition. \textbf{Bardin:} Conceptualization, Methodology, Writing - Original Draft, Writing - Review and Editing, Data Curation, Software.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\begin{landscape}
\begin{table}[h]
    \centering
    \begin{threeparttable}
    \caption{Descriptive summary of survey results}
    \begin{tabular}{l c c c c c c c c c c c c}
         \hline
                    & \multicolumn{4}{1}{Subfield}  & & \multicolumn{3}{1}{Approach} & &         &   & \\
         Measure    & PH & MT & NS & HU             & & QN & MX & QL                 & & Overall & N & Missing \\
         \hline
         Definition & \begin{tabular}[c]{@{}c@{}}1.73\\ (1.11)\end{tabular} &
                        \begin{tabular}[c]{@{}c@{}}2.00\\ (1.09)\end{tabular} & 
                        \begin{tabular}[c]{@{}c@{}}1.86\\ (1.06)\end{tabular} &
                        \begin{tabular}[c]{@{}c@{}}1.84\\ (1.14)\end{tabular} & 
                        & 
                        \begin{tabular}[c]{@{}c@{}}1.91\\ (1.13)\end{tabular} & 
                        \begin{tabular}[c]{@{}c@{}}1.82\\ (1.04)\end{tabular} &
                        \begin{tabular}[c]{@{}c@{}}1.69\\ (1.26)\end{tabular} & 
                        & 
                        \begin{tabular}[c]{@{}c@{}}1.83\\ (1.12)\end{tabular} &
                        181 &
                        37 \\
         Familiarity &  \begin{tabular}[c]{@{}c@{}}3.71\\ (0.98)\end{tabular} & 
                        \begin{tabular}[c]{@{}c@{}}3.97\\ (0.86)\end{tabular} & 
                        \begin{tabular}[c]{@{}c@{}}2.82\\ (1.76)\end{tabular} &
                        \begin{tabular}[c]{@{}c@{}}2.36\\ (1.36)\end{tabular} &
                        &
                        \begin{tabular}[c]{@{}c@{}}3.75\\ (0.98)\end{tabular} &
                        \begin{tabular}[c]{@{}c@{}}3.46\\ (1.22)\end{tabular} &
                        \begin{tabular}[c]{@{}c@{}}1.75\\ (1.30)\end{tabular} &
                        &
                        \begin{tabular}[c]{@{}c@{}}3.26\\ (1.36)\end{tabular} &
                        218 &
                        0 \\
         Experience & \begin{tabular}[c]{@{}c@{}}2.06\\ (1.18)\end{tabular} &
                        \begin{tabular}[c]{@{}c@{}}1.85\\ (1.38)\end{tabular} & 
                        \begin{tabular}[c]{@{}c@{}}1.18\\ (0.96)\end{tabular} &
                        \begin{tabular}[c]{@{}c@{}}0.60\\ (0.80)\end{tabular} &
                        &
                        \begin{tabular}[c]{@{}c@{}}1.63\\ (1.32)\end{tabular} &
                        \begin{tabular}[c]{@{}c@{}}1.69\\ (1.24)\end{tabular} &
                        \begin{tabular}[c]{@{}c@{}}0.48\\ (0.68)\end{tabular} &
                        &
                        \begin{tabular}[c]{@{}c@{}}1.44\\ (1.28)\end{tabular}
                        &
                        218 &
                        0 \\         
         Barriers & \begin{tabular}[c]{@{}c@{}}8.31\\ (3.00)\end{tabular} &
                        \begin{tabular}[c]{@{}c@{}}9.31\\ (2.79)\end{tabular} & 
                        \begin{tabular}[c]{@{}c@{}}7.44\\ (3.20)\end{tabular} &
                        \begin{tabular}[c]{@{}c@{}}7.32\\ (3.57)\end{tabular} &
                        &
                        \begin{tabular}[c]{@{}c@{}}8.78\\ (2.80)\end{tabular} &
                        \begin{tabular}[c]{@{}c@{}}8.58\\ (2.96)\end{tabular} &
                        \begin{tabular}[c]{@{}c@{}}5.97\\ (3.83)\end{tabular} &
                        &
                        \begin{tabular}[c]{@{}c@{}}8.20\\ (3.24)\end{tabular} &
                        182 &
                        36 \\
        \hline
    \end{tabular}
    \begin{tablenotes}
        \footnotesize
        \item Acronyms indicate: \textit{PH} Physical Geography, \textit{MT} GIScience and Methods, \textit{NS} Nature and Society, \textit{HU} Human Geography; \textit{QN} Quantitative, \textit{MX} Mixed Methods, \textit{QL} Qualitative. 
    \end{tablenotes}
    \label{tab:overview}
    \end{threeparttable}
\end{table}
\end{landscape}

%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\begin{landscape}
\begin{table}[h]
    \centering
    \begin{threeparttable}
    \caption{Barriers to Reproducibility}
    \begin{tabular}{l c c c c c c c c c c}
         \hline
                    & \multicolumn{4}{1}{Subfield}  & & \multicolumn{3}{1}{Approach} & & \\
         Barrier    & PH & MT & NS & HU            & & QN & MX & QL              & & Overall \\
         \hline
         \textit{Research Environment}      & & & & & & & & & & \\
         Pressure to Publish                & 83.1\% & 75.0\% & 54.5\% & 61.2\% & & 77.3\% & 75.3\% & 47.5\% & & 71.5\% \\
         Insufficient oversight             & 81.6\% & 82.0\% & 63.6\% & 56.7\% & & 82.6\% & 74.2\% & 40.0\% & & 71.1\% \\
         Fraud                              & 32.3\% & 39.3\% & 9.1\%  & 22.4\% & & 29.4\% & 31.8\% & 20.0\% & & 28.4\% \\
                                            & & & & & & & & & &\\
         \textit{Artifact Availability}     & & & & & & & & & & \\
         Insufficient metadata              & 76.9\% & 78.7\% & 59.1\% & 62.6\% & & 79.4\% & 72.9\% & 45.0\% & & 80.2\% \\
         Unavailable Data                   & 89.3\% & 78.6\% & 59.1\% & 65.7\% & & 87.0\% & 81.2\% & 42.5\% & & 75.2\% \\
         Unavailable protocol/code          & 77.0\% & 86.9\% & 50.0\% & 59.7\% & & 87.0\% & 70.6\% & 37.5\% & & 71.1\% \\
         Published incomplete results       & 73.8\% & 78.7\% & 63.7\% & 56.7\% & & 78.2\% & 72.9\% & 37.2\% & & 68.4\% \\
         Use restricted data/software       & 56.9\% & 82.0\% & 40.9\% & 44.8\% & & 70.7\% & 56.5\% & 32.5\% & & 57.8\% \\
                                            & & & & & & & & & &\\
         \textit{Study Characteristics}     & & & & & & & & & & \\
         Geographic variation               & 80.0\% & 72.2\% & 81.8\% & 61.2\% & & 79.4\% & 74.1\% & 40.0\% &  & 71.5\% \\
         Researcher positionality           & 50.8\% & 70.5\% & 72.8\% & 70.2\% & & 56.5\% & 65.9\% & 80.0\% &  & 64.2\% \\
         Chance                             & 66.2\% & 65.6\% & 68.2\% & 55.2\% & & 67.4\% & 64.4\% & 52.5\% &  & 62.3\% \\
         Different computation              & 53.8\% & 68.9\% & 31.8\% & 38.8\% & & 63.1\% & 56.5\% & 25.0\% &  & 50.9\% \\
                                            & & & & & & & & & &\\
         N                                  & 65 & 61 & 22 & 67 & & 92 & 85 & 40 & & 218 \\
        \hline
    \end{tabular}
    \begin{tablenotes}
        \footnotesize
        \item Cells report the percentage of respondents reporting each factor occasionally or frequently contributed to a lack of reproducibility in geographic research. Acronyms indicate: \textit{PH} Physical Geography, \textit{MT} GIScience and Methods, \textit{NS} Nature and Society, \textit{HU} Human Geography; \textit{QN} Quantitative, \textit{MX} Mixed Methods, \textit{QL} Qualitative. 
    \end{tablenotes}
    \label{tab:barriers}
    \end{threeparttable}
\end{table}
\end{landscape}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\noindent PETER KEDRON is an Associate Professor in the School of Geographical Science and Urban Planning and core faculty member in the Spatial Analysis Research Center (SPARC) at Arizona State University, Tempe, AZ, 85283, US. Email: Peter.Kedron@asu.edu. His research interests include spatial analysis, geographic information science, economic geography, and the accumulation of knowledge about geographic phenomena. \\

\noindent JOSEPH HOLLER is an Assistant Professor of Geography at Middlebury College, Middlebury, VT, 05753, US. Email: \\

\noindent SARAH BARDIN is a PhD candidate ...

\end{document}