### 2023-05-31 docs/manuscript/saved_text.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Assessing the Reproducibility of Research}

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{The Evidence Currently Available from Reproducibility Assessments}
The reproducibility of research can be assessed directly by attempting to recreate the results of selected studies, or indirectly by reviewing the availability of the research components (e.g., data and code) or surveying those working in that field about their research practices. 
Direct assessments of reproducibility provide insight into how a particular study was conducted and why a consistent result could, or could not, be obtained.  
However, such assessments are resource and time intensive, which makes it difficult to gather information from numerous studies. 
Constrained to a limited number of studies, direct assessments offer only suggestive evidence about the reproducibility of unevaluated studies and the prevalence of reproducible research practices. 
In contrast, comparatively less resource intensive indirect assessments that catalog reproducible research practices across many studies offer less insight into individual studies, but provide evidence of how common reproducible research practices are and, by extension, how reproducible work in a field of study is likely to be.

Across many disciplines, direct assessments of reproducibility remain relatively rare and attempted reproductions in the biomedical sciences \citep{iqbal2016reproducible}, economics \citep{chang2015economics}, computer science \citep{moraila2014measuring}, and psychology \citep{open2015estimating} have all been able to recreate only a portion of results. 
Within geography, the small number of available reproduction attempts similarly suggest that many studies cannot be fully reproduced, or are simply missing the components needed to attempt a reproduction \citep{nust2018, Kedron2021ssrn, konkol2019, ostermann2021, Nust_AGILE_2022}.
For example, \citet{ostermann2021}'s attempt to reproduce the computational analyses of 75 papers published in the GIScience conference series found that most results could not be recreated, or could only be recreated with significant effort. 
\citet{konkol2019}'s attempt to reproduce 41 geocomputational studies likewise found that 39 of the papers had coding issues that hindered reproduction and that 47 percent of the figures created using the original data and code deviated from the those published in some significant way (e.g., graphs with different curves). 
A series of reproduction attempts by \citet{Kedron_MollaloRP, Kedron_SaffaryRP, Kedron_VijayanRP} and \citet{paez2022reproducibility} show that the issues hindering the reproduction of results extend beyond computation and into the conceptualization and design of geographic research.
While indicative, these studies provide only suggestive evidence of underlying problems, not a comprehensive assessment of the prevalence or sources of these issues.

Indirect reproducibility assessments offer a wider view of what and how common these underlying issues may be. 
The available evidence from reviews of studies across the sciences suggests inadequate record keeping, opaque reporting, unavailable research components, and a lack of incentives are likely important factors contributing to the irreproducibility of research \citep{NASEM2019}. 
Systematic reviews of publications in prominent outlets \citep{byrne_2017,stodden2018empirical} suggest these issues are likely widespread. 
Reproducibility surveys paint a similar picture. 
Recent surveys suggest that the majority of scientists have either not tried, or tried and failed to reproduce another scientist's work \citep{baker20161, boulbes2018survey}. 
Moreover, a sizable portion of the researchers responding to these surveys admit to engaging in questionable research practices linked to the publication of false positive results and low reproducibility rates \citep{fanelli2009many, fraser2018questionable}.
These findings match researcher perceptions, which link irreproducibility to causes ranging from natural variation to outright fraud \citep{ranstam2000fraud, anderson2007normative, baker20161}.

To date only a handful studies have attempted to indirectly measure the reproducibility of geographic research.
In a survey of remote sensing editors and working group officers of the International Society for Photogrammetry and Remote Sensing, \citet{balz2020reproducibility} asked respondents whether they believe a reproducibility crises existed in their sub-field, but not what the causes of such a crisis might be.
In a richer treatment of the issue, \citet{konkol2019} surveyed of participants from the 2016 European Geosciences Union General Assembly to assess the frequency with which researchers published their work in ways that enabled computational reproducibility. 
Those authors found that only 33 percent of respondents included links to the data used in their analyses and only 12 percent provided their analytical code. 
The authors also found that only seven percent of respondents ever attempted to reproduce the work of other researchers.
Similarly, in a literature review of volunteered geographic information research, \citet{ostermann2017} found the computational reproducibility of work in the sub-field limited by the availability of original data, metadata, source code, or pseudocode.

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Limitations of the Currently Available Evidence}
Despite their growing number, the available body of reproducibility assessments currently provides insufficient evidence to conclusively evaluate the reproducibility of research generally, or disciplinary research specifically \citep{NASEM2019}.
This knowledge gap exists because neither direct nor indirect reproducibility assessments have been designed so the results obtained from the sample of studies evaluated can be extended to the population of research as a whole.
Most reproducibility reviews sample studies from conveniently accessible archives, such as conference paper series \citep{gundersen2018state, ostermann2017}, or target specific journals and disciplinary repositories \citep{stodden2018empirical, stodden2018enabling, byrne_2017}.
These assessments have also tended to focus on evaluating the computational components of research such as data and code availability, as this information is easier to identify and assess at scale when compared to components like field procedures or qualitative analysis that may be included only in text.

Surveys of researcher practices and attitudes on reproducibility face similar challenges. 
Rather than carefully defining researcher populations of interest, existing surveys have drawn data non-randomly from self-selecting populations that are convenient to survey (e.g., members of professional association, conference attendees), but are not necessarily representative of researchers active in a particular field. 
These surveys have also commonly failed to systematically report the methodological details (e.g., response rate) needed to assess and address potential bias in survey response.

% Some saved language about survey biases from limitations (from Peter)
Without this knowledge and data, we have no way of knowing if an adjustment to our sample is likely to increase or decrease bias in our results.

Being unable to meaningfully differentiate subgroups for either stratification or post-stratification, we adopted a simple random sampling scheme and provide our data to facilitate transparency and reanalysis of our sample.
Considering these challenges, our results could be viewed as an exploratory analysis that identifies researcher characteristics that might differentiate subgroups with similar understandings of reproducibility, which could be used to stratify or post-stratify future reproducibility related research in geography.
Nonetheless, because we were unable to assess the representativeness of our sample our results may be subject to at least three forms of bias.

\textit{Exclusion Bias:} One limitation of our approach is that geographic researchers publish in a range of journals that are not necessarily indexed as geography by the Web of Science. 
Geographic researchers that have only published outside these indexed journals would not be captured in our sampling frame creating non-coverage errors in our sample. 
However, we believe the number of individuals falling into this category will be small as most active geographic researchers are likely to publish at least one study in one of the X geography journals used in our sampling frame over a 5-year period.

\textit{Self-selection Bias}. A related concern is that it may be the case that geographic researchers more familiar with reproducible research practices, or those working in subfields more involved with the current reproduciblity debate (e.g., quantitative, computational research) were more likely to respond to our survey. 
Conversely, it may also be the case that researchers working in subfields traditionally associated with critiques of a positivist scientific approach (e.g., qualitative, human geography) were less likely to participate in our survey. 
We attempted to partially quantify this issue by examining the characteristics of the 40 researchers who started our survey but did not complete enough of the survey to be included in the final analysis.
Taking into account 29 researchers who reported their subfield but did not complete the full survey, completion rates for all subfields were between 84 and 87 percent, except slightly higher rates for methods / GIScience (96.8\%).
Considering 31 researchers who reported their method but did not complete the full survey, completion rates were 84.2 percent for mixed methods, 87.0 percent for qualitative methods, and 91.1 percent for quantitative methods. 
These values suggest that self-selection was not a significant issue.
This analysis of course does not measure researchers who elected to never start the survey.

While we cannot definitively eliminate the possibility of self-selection bias from our survey, we believe its presence may be a limited concern because it is likely to only amplify some of our key findings. 
Namely, that reproducible research practices have yet to be widely adopted in the field and that relatively few geographers are attempting and sharing independent reproduction studies.
Nonetheless, we believe concerns about self-selection should motivate a closer examination of reproducibility within the subfields and approaches for which we obtained the smallest samples --- qualitative, human geography, and nature society.

\textit{Questionnare Bias:} Another potential limitation of our research is that we based our survey instrument on a review of prior reproducibility surveys and our identification of practices and barriers discussed in the reproducibility literature. 
To the extent that this literature over represents perspectives from selected disciplines, experimental research designs, and computationally-intensive research approaches our survey may reflect this bias. 
If this were the case, our survey may not gather data on practices and barriers important to the reproducibility of types of geographic research not like the work conducted in those disciplines that are well represented in the existing literature. 
To address this concern, we incorporated into our survey instrument questions informed by a parallel review of the reproducibility literature available within geography and a review of critiques of positivist science made by social scientists and human geographers.
We also provided the option for an open ended text response to questions about reproducible research practices, researcher familiarity with those practices, and barriers to reproducibility in an attempt to identify factor and issues we did not anticipate during instrument construction.