### 2023-07-14 docs/presentation/preprocess.Rmd
viz: refactor constants

---
title: "Preprocessing"
output: 
  html_document:
    theme: flatly
    df_print: paged
date: "2023-07-12"
---

Making sense of the survey data.

The file `analysis_hegs_rpr.csv` was saved from `data\derived\public\analysis_hegs_rpr.rds`.

```{r import, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  fig.retina = 4,
  fig.showtext = TRUE,
  message = FALSE,
  warning = FALSE
)

# all your themes are belong to us
# library(ggthemes)
library(readr)
library(tidyverse)
library(magrittr)
library(showtext)
font_add_google("Fira Sans", "Fira Sans")
```

```{r data}
analysis <- read_csv("data/analysis_hegs_rpr.csv", col_select = starts_with("Q"))
source("data/constants.R")
```

## Groups

Dropping NAs for brevity

```{r drop na}
analysis %<>%
  drop_na(Q3_recoded, Q4) %>%
  filter(Q3_recoded != "Other")
```

Question 3: primary subfield of research

```{r q3 groups}
analysis %>%
  group_by(Q3_recoded) %>%
  summarise(n = n())
```

Question 4: type of research

```{r q4 groups}
analysis %>%
  group_by(Q4) %>%
  summarise(n = n())
```

Q19: demographics

Potentially split into professor - not professor

```{r q19 groups}
analysis %>%
  group_by(Q19_recoded) %>%
  summarise(n = n())
```

## Word cloud

Q6
Q10
Q12a
Q15
Q17a

Potentially per group, but only for Q6

```{r DRY wordcloud}
column_to_cloud <- function(
    col, nrows = 30,
    data = analysis,
    my_palette = c(
      "#cdb4db",
      "#6d597a",
      "#b56576",
      "#e56b6f",
      "#eaac8b"
    )) {
  require(tidytext)
  word_counts <- data %>%
    drop_na(all_of(col)) %>%
    unnest_tokens("word", col) %>%
    # https://bookdown.org/psonkin18/berkshire/tokenize.html#text-normalization---stemming-and-lemmatization
    # doesn't work - why is `reproduct` different from `reproduc`?
    # mutate(word = SnowballC::wordStem(word, language = "english")) %>%
    anti_join(stop_words, by = "word") %>%
    count(word) %>%
    arrange(desc(n)) %>%
    slice(1:nrows)

  require(wordcloud2)
  wordcloud2(
    word_counts,
    color = rep_len(
      my_palette,
      nrow(word_counts)
    )
  ) %>%
    return()
}
```

```{r word clouds}
clouds <- map(cloud_cols, column_to_cloud) %>%
  set_names(cloud_cols)
# saveRDS(clouds, file = "data/clouds.rds")
```

```{r q6 cloud}
clouds_q6_q3 <- map(
  subfields, ~ analysis %>%
    filter(Q3_recoded == .x) %>%
    column_to_cloud("Q6", data = .)
) %>%
  set_names(subfields)

clouds_q6_q4 <- map(
  types, ~ analysis %>%
    filter(Q4 == .x) %>%
    column_to_cloud("Q6", data = .)
) %>%
  set_names(types)

clouds_q6 <- append(clouds_q6_q3, clouds_q6_q4)
clouds_q6$Everyone <- clouds$Q6
# saveRDS(clouds_q6, file = "data/clouds_q6.rds")
```

## Density plots
Include animations?

Q5

```{r q5}
analysis %>%
  mutate(Q5_num = as.numeric(four_levels[Q5])) %>%
  ggplot() +
  scale_x_continuous(
    breaks = as.numeric(four_levels),
    labels = str_wrap(names(four_levels), width = 9)
  ) +
  geom_density(aes(x = Q5_num), adjust = 1.6) +
  theme_minimal() +
  theme(
    text = element_text(family = "Fira Sans", size = 16),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  labs(
    title = "How familiar are you with the term \"reproducibility\"?",
    subtitle = "In the context of conducting research",
    x = "", y = ""
  )
```

Q7 (many parts)

Q13 (percentages)

## Likert charts
Q8

Q9 (maybe)

Q11
Q12

Q14

Q17

## Bar plot
Or something simple
Include animations

Q16

## Demographics